---
output:
  html_document: default
  pdf_document: default
---
Installing the packages: 

```{r}

#install.packages("ggplot2")
#install.packages("palmerpenguins")
#install.packages("janitor")
#install.packages("dplyr")
#install.packages("tinytex")

```

Loading the packages: 

```{r}
library(ggplot2)
library(palmerpenguins)
library(janitor)
library(dplyr)
library(tinytex)

```
Sourcing my functions: 

```{r}
source("Functions/Cleaning.functions.R")
source("Functions/Graph.function.R")
source("Functions/plotting.R")

```


Observing the raw data: 

```{r}


head(penguins_raw)
```

Blotting a bad graph with the raw data 

Plotting a bad graph: 

```{r, echo=FALSE}


write.csv(penguins_raw, "Data/penguins_raw.csv")

penguins_clean <- penguins_raw %>%
    clean_function() %>%
    shorten_species() %>%
    remove_empty_columns_rows()


#Saving my cleaned data: 

write.csv(penguins_clean, "Data/penguins_clean.csv")



# Define a custom color scale with very similar shades of red and orange
my_colors <- c("#FF7F50", "#FF6347")

penguin_barplot <- ggplot(data = penguins_clean, aes(x = sex, y = body_mass_g, fill = sex)) + 
  geom_bar(stat = "identity", position = "dodge") + 
  scale_fill_manual(values = my_colors) +  # Use the custom color scale
  labs(title = "", x = "", y = "Body Weight") + 
  theme_minimal() + 
  coord_cartesian(ylim = c(3000, 6500)) +
  theme(panel.grid = element_blank()) +
  guides(fill = FALSE)

# Print the plot
print(penguin_barplot)

```





I have drawn a cumulative bar graph which shows differences in body mass between male and female penguins by adding together the total body masses of all the penguins. Firstly, there is no title to my graph so there is no indication of what relationship between my two parameters my graph is showing. There is no indication that the data is about penguins and relating to the differences in body mass between the different sexes.  There is no title to my x axis, besides ‘male’ and ‘female,’ so the reader must infer that I am investigating difference in body mass between sexes. The title to my y axis ‘Body mass’ does not give a unit so there is no indication on how large the differences between the sexes are. I have also removed the grid lines from my background making it more difficult to assess how large the differences are. I have not calculated standard deviation or made these into error bars. I am therefore hiding elements of my data which may not be statistically significant. I have also not excluded NA values where an individuals sex was not recorded, but this has been included in my graph. This is irrelevant data and distracts readers from the main findings. I have also not included a figure legend, making it more difficult for the reader to realise which column applies to each sex. The colours of the male and female bars are very similar making them difficult to distinguish. They are also not color blind friendly, presenting accessibility issues. Finally, I have truncated the y axis so that it does not start at 0 but starts at 3000. In a study where participants observed both truncated and non truncated graphs, the differences in truncated graphs were found to be perceived as larger than the control graphs with a statistically significant difference. 

References: 
Yang, B. W., Vargas-Restrepo, C., Stanley, M. L., & Marsh, E. J. The Misleading Effect of Truncating Bar Graphs.

Introduction: 

My analysis will investigate how body mass in the palmer penguins varies with flipper length. The palmer penguins species are the Chinstrap species, the Gentoo penguins, and the Adelie penguins. I am investigating to see if there might be a correlation between increasing flipper length and body mass. My hypothesis is that as flipper length increases, body mass will increase, as a consequence of the general morphology and dimensions of the birds remaining the same as flipper length increases (instead of perhaps, torso size decreasing). I will use a linear regression model to find out if there is a positive linear relationship between increasing flipper length and body mass. 

Cleaning my data: 

First I must save the raw data: 


```{r}
write.csv(penguins_raw, "Data/penguins_raw.csv")

```

Cleaning my data: 

```{r}
penguins_clean <- penguins_raw %>%
    clean_function() %>%
    shorten_species() %>%
    remove_empty_columns_rows()

#The above are functions that I made to clean the data. I have stored the functions in a folder in my repository called Functions. The file is called Cleaning.Functions.R, where I have explained the purpose of each function. 


#Saving my cleaned data: 

write.csv(penguins_clean, "Data/penguins_clean.csv")


```

```{r}
print(penguins_clean)

```

Exploratory plot: 

```{r}
#I made a function which makes my ggplot scatter plot. The function can be found in the file called Graph.function.R inside my functions folder in my repository. 
penguin.scatterplot <- scatter.plot.flippers(penguins_clean)
penguin.scatterplot

```
Size, scale and resolution: 
```{r}
#If the "ragg" package is not installed, install this package: 
#install.packages("ragg")
library(ragg)


```

Saving the png file: 

```{r}

save_scatter_plot_png(penguins_clean,
                      "figures/scatter.png",
                      size = 500, res = 300, scaling=0.25)
```

Creating vector images: 

The image above is problematic because when you zoom in they are blurry because they are created by a matrix of pixels. 

Saving my vector image: 

```{r}
#If the svg package is not installed: 
library(svglite)
#Saving the plot 
save_flipper_plot_svg(penguins_clean, 
                      "figures/scatter_vector.svg", 
                      size = 15, scaling = 1)

```



Statistical tests: 



```{r Statistics}

#Creating a linear model for my two variables: 

linear_model <- lm(body_mass_g ~ flipper_length_mm, data = penguins_clean)

#Displaying the summary of the linear model: 

summary(linear_model)

```

### Results & Discussion

There is a positive linear relationship between flipper length in and body mass.
The adjusted R squared value is 0.7583, which means that 75.8% of the variation in body mass can be explained by increasing flipper length. The p value is 2.2 x 10^-16. This value is < 0.05. This means that the linear relationship between flipper length and body mass is statistically significant. The gradient is 49.686 so for every mm that flipper length increases, body mass increases by 49.5g. 

Link to my repository: 

https://github.com/Mynewgithubabc/Penguin.Assignment.git

Link to my partner's repository: 

https://github.com/palmpeng1/assignment

My experience of running my partner’s code: 


Running my partners code was straightforward as it is reproducible, meaning that other readers can easily understand and run the code. They have used piping to carry out multiple steps on the data frame at once without repeating multiple lines of code, making the code easier for me to understand. The piping step includes multiple new functions which my partner has created called clean_columns, clean_species etc that clean the data. They have explained before their piping code that the functions are stored in a separate file called cleaning.R instead of printing the entire code to make the functions in this R markdown. They have instead only called the functions in the pipe. Between each piping step my partner has annotated each step in the cleaning pipeline so it is easy to understand what each function in the pipeline is doing. 
My partner also created a function called plot_function which they have explained is stored in a plotting.R file. My partner only calls this function in their R markdown, avoiding repetition of many lines of code. 

Keeping functions in a separate plotting.r and cleaning.r instead of writing the code in the note book and repeating it multiple times on different data sets is important as it keeps a safe copy of the original code that is version controlled. This allows you to revert any changes that were made to the graph or data. This is particularly important for graph functions as it allows you to make the exact same figure again, multiple times, on different datasets. Creating functions prevents the need to write the same lines of code out repeatedly making code easier to understand. 

However, although my partner has explained that they have stored the functions in separate files, they have not explained the fact that they have made new functions (called clean_columns(), shortern_species(), etc that clean the data. Therefore if a new reader is not familiar with creating functions, they may not be able to follow the pipeline easily, or trouble shoot anything not working.

I did not encounter any errors in running my partners code
However, since my partner has stored their functions in the separate folder called functions, and in a separate file called cleaning.R and plotting,r, it would be easy to navigate to find the functions and source any errors in their code. 

Reflecting on my code based on my experience of my partner’s code and their review of my code: 

•	My partner suggested that I explain what the cleaning loops do in the chunk itself. I noted that they had done this in their code too. I agree with this, as it would help the reader to understand what my cleaning functions are doing without having to check the cleaning.r file for the explanation. This would make my code more human readable. My partner also suggested I create an extra file storing the package versions used in the analysis. I agree with this because it would allow the same analysis to be run in the future even if the packages have been updated, increasing reproducibility in the future. 
My partner suggested that I could use an ‘if’ loop to only install packages if they are not already installed. I agree with this as it would save time and computer energy. 
My partner noted that my file names for my functions are slightly confusing. My plotting.r is used for saving my plots, whereas my graph.functions.r is used for saving my graph functions. I have now realised that this is not human readable. 

What I learnt about writing code for other people: 

I found that my partner’s code was easier to interpret than my own. The biggest difference between mine and my partner’s code was that my partner used comments all the way through their code explaining what each step was doing. For example they stated where their functions were stored (for example in plotting.r), and they explained what each function in the pipe was doing. This was different to my code.My main notebook simply contains lines of code, without explaining what each line is doing. Although in my functions folder, (eg cleaning.r) I have explained what all my functions are doing, I find my partner’s version more human readable and therefore more reproduceable. 
I have also not included any explanation, analysis, or justification as to why I used a linear model to investigate the relationship between my two parameters, flipper length and body mass. This makes my analysis less human readable as it is difficult for another reader to grasp what calculations I am using and why. 
I have learnt that for other people to use the code it must be human readable. This means that the notebook is as simple as possible, with as few as possible lines of code. This was achieved for example by making functions and storing the lines of code to do this in separate functions such as cleaning.r.

If I was to complete this task again I would include annotations within the code explaining that these functions have been made, what they do, and where they are stored. This makes the code easy to interpret, so it is both computer and human readable. 



